---
title: "index"
author: "Pietro Gobbi"
date: "10/20/2020"
output: html_document
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=8, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyquant)
```

```{r get_rates, warning=FALSE}
# Get a list of FRED codes for US rates and US yield curve; choose monthly frequency
# to see, eg., the 3-month T-bill https://fred.stlouisfed.org/series/TB3MS
tickers <- c('TB3MS', # 3-month Treasury bill (or T-bill)
             'TB6MS', # 6-month
             'GS1',   # 1-year
             'GS2',   # 2-year, etc....
             'GS3',
             'GS5',
             'GS7',
             'GS10',
             'GS20',
             'GS30')  #.... all the way to the 30-year rate

# Turn  FRED codes to human readable variables
myvars <- c('3-Month Treasury Bill',
            '6-Month Treasury Bill',
            '1-Year Treasury Rate',
            '2-Year Treasury Rate',
            '3-Year Treasury Rate',
            '5-Year Treasury Rate',
            '7-Year Treasury Rate',
            '10-Year Treasury Rate',
            '20-Year Treasury Rate',
            '30-Year Treasury Rate')

maturity <- c('3m', '6m', '1y', '2y','3y','5y','7y','10y','20y','30y')

# by default R will sort these maturities alphabetically; but since we want
# to keep them in that exact order, we recast maturity as a factor 
# or categorical variable, with the levels defined as we want
maturity <- factor(maturity, levels = maturity)

# Create a lookup dataset
mylookup<-data.frame(symbol=tickers,var=myvars, maturity=maturity)
# Take a look:
mylookup %>% 
  knitr::kable()

df <- tickers %>% tidyquant::tq_get(get="economic.data", 
                   from="1960-01-01")   # start from January 1960

glimpse(df)
```

Our dataframe `df` has three columns (variables):

- `symbol`: the FRED database ticker symbol
- `date`: already a date object
- `price`: the actual yield on that date

The first thing would be to join this dataframe `df` with the dataframe `mylookup` so we have a more readable version of maturities, durations, etc.


```{r join_data, warning=FALSE}

yield_curve <-left_join(df,mylookup,by="symbol") 

```

```{r setup_US-recessions, warning=FALSE}

# get US recession dates after 1946 from Wikipedia 
# https://en.wikipedia.org/wiki/List_of_recessions_in_the_United_States

recessions <- tibble(
  from = c("1948-11-01", "1953-07-01", "1957-08-01", "1960-04-01", "1969-12-01", "1973-11-01", "1980-01-01","1981-07-01", "1990-07-01", "2001-03-01", "2007-12-01"),  
  to = c("1949-10-01", "1954-05-01", "1958-04-01", "1961-02-01", "1970-11-01", "1975-03-01", "1980-07-01", "1982-11-01", "1991-03-01", "2001-11-01", "2009-06-01") 
  )  %>% 
  mutate(From = ymd(from), 
         To=ymd(to),
         duration_days = To-From)

recessions
```

```{r}

threemonths <- yield_curve %>%
  filter(maturity=="3m")%>%
  select(date,price)
colnames(threemonths) <- c("date","priceTBills")

tenyears <- yield_curve %>%
  filter(maturity=="10y")%>%
  select(date,price)
colnames(tenyears) <- c("date","priceTeanYear")


diff <- merge(threemonths,tenyears)
diff <- diff %>%
  mutate(spread=priceTeanYear-priceTBills, minus_diff = ifelse(spread >= 0, spread, 0), plus_diff = ifelse(spread<0, spread, 0))

recessions$to <- as.Date(recessions$to)
recessions$from <- as.Date(recessions$from)
recessions_graph <- recessions %>%
  filter(to >= "1959/01/01")

ggplot(diff,aes(x=date, y=spread))+
  geom_line()+
  geom_hline(yintercept=0)+
  labs(title= "Yield Curve Inversion: 10-year minus 3-months US Treasury rates", subtitle = "Difference in % points, monthly averages.\nShaded areas correspond to recessions", caption = "Source: St. Louis Federeal Reserve Economic Database (FRED)", x="", y="Difference ( 10 year-3 month ) yield in %")+
  geom_rug(aes(colour=ifelse(spread>=0,">=0","<0")),sides="b")+
  scale_colour_manual(values=c("#eab5b7","#abc7e6"), guide=FALSE)+
  geom_ribbon(aes(ymin=0,ymax=minus_diff),fill="#abc7e6",alpha=0.6)+
  geom_ribbon(aes(ymin=0,ymax=plus_diff),fill="#eab5b7",alpha=0.6)+
  theme_minimal()+
  geom_rect(data=recessions_graph, inherit.aes=FALSE,
            aes(xmin=from,xmax=to,ymin=-Inf,ymax=Inf, alpha=0.2))+
  scale_x_date( breaks ="2 years",  date_labels = "%Y")+
  theme(legend.position = "none", text = element_text(size=7.3), plot.title = element_text(face="bold"))
```
## Plotting the yield curve


### Yields on US rates by duration since 1960


```{r}
yield_curve$var = factor(yield_curve$var, myvars)

ggplot(yield_curve, aes(x=date , y = price, color = var))+
  geom_line()+
  theme_bw()+
  facet_wrap(~var, ncol = 2)+
  guides(color = FALSE)+
  labs(title= "Yields on U.S. Treasuries rates since 1960", x="", y="%", caption = "Source: St. Louis Federeal Reserve Economic Database (FRED)")+
  theme(text = element_text(size=8))
  

```

### Monthly yields on US rates by duration since 1999 on a year-by-year basis



```{r}
graph2 <- yield_curve %>%
  filter(date>=as.Date("1999/01/01"))%>%
  mutate(year = format(date, "%Y"), month = format(date, "%m"))


ggplot(graph2, aes(x=maturity, group = month, color = year))+
  geom_line(aes(y=price))+
  theme_bw()+
  facet_wrap(~year, ncol = 4)+
  labs(title = "US Yield Curve", caption =  "Source: St. Louis Federeal Reserve Economic Database (FRED)", x="Maturity" , y="Yield (%)")+
  guides(color=FALSE)+
  theme(text = element_text(size = 9))

```


### 3-month and 10-year yields since 1999

```{r}
graph3 <- yield_curve%>%
  filter(date>=as.Date("1999/01/01"), maturity == "10y"|maturity == "3m")

ggplot(graph3 , aes(x= date, y=price, color=var))+
  geom_line()+
  theme_bw()+
  labs(title= "Yields on 3-month and 10-year US Treasury rates since 1999", caption = "Source: St. Louis Federeal Reserve Economic Database (FRED)", x="", y="%")+
  theme(legend.title = element_blank(), text=element_text(size=8))


```